### 2019-08-07 acknowledgements.tex
submission files

\documentclass{article}
\usepackage{fullpage}
\author{Acknowledgements}
\title{Case study validations of automatic bullet matching}
\date{}
\begin{document}
\maketitle

This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.\vspace{1cm}

The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us. 
\end{document}