### 2019-08-07 cover_letter.tex

\documentclass[11pt]{article}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{float}
% \setlength{\textheight}{26cm}
% \setlength{\textwidth}{18cm}
\topmargin0in

\oddsidemargin-1cm
\evensidemargin0in
\textheight9in
\textwidth7in
\setlength{\parindent}{0em}

\pagestyle{fancyplain}
\addtolength{\headheight}{\baselineskip}
% \addtolength{\headheight}{43pt}
\rhead{}
\cfoot{}
\markright{{\hfill\large \sc Center for Statistics and Applications in Forensic Evidence}\\{\hfill\sc Iowa State University} \\{\hfill 195 Durham Center, 613 Morill Rd., Ames, IA 50011}}

\begin{document}
C. Jackowski \\
Institute of Forensic Medicine\vspace{.5em}\\
C. Cattaneo \\
Università degli Studi di Milano \vspace{2em}

\today\\\vspace{1em}

{\bf Re: Submission to Forensic Science International}\\\vspace{2em}

Dear Professor Jackowski and Professor Cattaneo, \vspace{1em}

Please find attached a manuscript, ``Case study validations of automatic bullet matching" intended for submission to FSI. \vspace{1em}

The manuscript describes a validation of an automatic bullet matching algorithm on several sets of bullets used to test examiner proficiency. This algorithm uses a scheme similar to that used by footwear forensic analysts, unlike other convolutional neural network based analyses in footwear forensics. We describe the model's structure, capabilities, and overall performance; we also explore situations where the model's automatic recognition process does not succeed and provide some possible solutions for these issues going forward. We believe this work lays the groundwork to move beyond theoretical applications of machine learning in toolmark forensics toward implementation of these algorithms in the crime lab. \vspace{1em}

I hope that you find the manuscript acceptable for publication in FSI. If there is any additional information I can provide, please feel free to contact us directly. \vspace{0.25in}

Sincerely,\\
\begin{minipage}{.25\linewidth}
\vspace*{.5cm}
\includegraphics[width=2.5in,keepaspectratio=TRUE]{signature}
\vspace*{.5cm}
\end{minipage}

Susan Vanderplas\\
Center for Statistics and Applications in Forensic Evidence\\
Iowa State University

\end{document}