### 2019-08-07 title_page.tex

\documentclass[titlepage]{article}
\title{Case study validations of automatic bullet matching}
\author{Susan Vanderplas and Heike Hofmann}
\date{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}

\begin{document}
\maketitle

\end{document}