### 2019-12-13 title_page.tex
Update authors on title page

\documentclass{article}
\usepackage{fullpage}

\title{Case study validations of automatic bullet matching}
\author{Susan Vanderplas$^\ast$$^1$, Melissa Nally$^2$, Tylor Klep$^3$, Cristina Cadevall$^4$, Heike Hofmann$^1$}

\date{$^\ast$ Corresponding author. email: srvander@iastate.edu\\[1em]
$^1$Center for Statistics and Applications in Forensic Evidence\\Iowa State University\\
195 Durham Center, 613 Morill Rd., Ames, IA 50011\\[1em]
$^2$Houston Forensic Science Center\\500 Jefferson St 13th Floor, Houston, TX 77002\\[1em]
$^3$Phoenix Police Department\\Laboratory Services Bureau\\621 W Washington St, Phoenix, AZ 85003\\[1em]
$^4$Sensofar Group\\Parc Audiovisual de Catalunya Ctra, BV-1274, Km 1, 08225 Terrassa, Barcelona, Spain}

\begin{document}
\maketitle

\end{document}