### 2019-12-24 title_page.tex

\documentclass{article}
\usepackage{fullpage}

\title{Comparison of three similarity scores for bullet LEA matching}
\author{Susan Vanderplas$^\ast$$^{1,2}$, Melissa Nally$^3$, Tylor Klep$^4$, Cristina Cadevall$^5$, Heike Hofmann$^1$}

\date{$^\ast$ Corresponding author. email: susan.vanderplas@unl.edu\\[1em]
$^1$Center for Statistics and Applications in Forensic Evidence, Iowa State University\\
195 Durham Center, 613 Morill Rd., Ames, IA 50011\\[1em]
$^2$Statistics Department, University of Nebraska Lincoln\\340 Hardin Hall North Wing, Lincoln, NE 68583-0963\\[1em]
$^2$Houston Forensic Science Center\\500 Jefferson St 13th Floor, Houston, TX 77002\\[1em]
$^3$Phoenix Police Department\\Laboratory Services Bureau\\621 W Washington St, Phoenix, AZ 85003\\[1em]
$^4$Sensofar Group\\Parc Audiovisual de Catalunya Ctra, BV-1274, Km 1, 08225 Terrassa, Barcelona, Spain}

\begin{document}
\maketitle

\end{document}